\LaTeX\ Author Guidelines for ICCV Proceedings

### 2021-03-13 chapters/3_algorithm.tex
Intregrate overleaf comments

\textbf{Linkage criteria and hierarchical trees} -- 
% We also denote as $S_u$ the cluster associated with node $u$ \TODO{Needed?}.
% We call two clusters $S,S'$ \emph{adjacent} if there exists at least one edge ${e_{ts}\in E}$ connecting a node $t\in S_u$ to a node $s\in S_v$. 
Let the interaction $\interact(S,S')$ between two clusters $S,S'$ be defined as a function $\interact{}:2^{V} \times 2^{V} \rightarrow \mathbb{R}$, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S$ and $S'$, i.e. ${(S \times S') \cap E}$. 
The linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
A \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} representing the merging order of an agglomerative algorithm, such that the leaves of the tree are in one-to-one correspondence with $V$. 
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor $(u \vee v)\in T$ of nodes $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
% \TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
% Algorithm \ref{main_alg} returns a binary tree denoted by $T^*$.
Given an agglomerative algorithm with merging tree $T$, let $h_T:T \rightarrow \mathbb{N}$ denote the \emph{dendrogram-height} of each $(u\vee v)\in T$, which is defined as the iteration number at which nodes $u,v\in V$ were merged by the algorithm (see example in Fig.~\ref{fig:intro_figure}a). We also define $\treeHeight(u,v)$ as the signed interaction $\interact{}(S,S')$ between the two clusters $S,S'$ that were merged at iteration $h_T(u\vee v)$: 
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}\label{eq:def_dendr_interact}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T_{\mathrm{R}}[u \vee v]), \text{\texttt{leaves}}(T_{\mathrm{L}}[u \vee v]) \big)
\end{equation}
% \OPTIONAL{In other words, $h_T(u,v)$ and $\treeHeight(u,v)$ represent the \emph{dendrogram-height} and \emph{signed-interaction} associated to each $u \vee v$ node in the tree $T$ (see toy example in \TODO{Fig.~1}).}

@@

In \emph{Phase~One}, \algname{} selects the pair of clusters with the highest absolute interaction $|\interact(S, S')|$, so that the most attractive and the most repulsive pairs are analyzed first. If the interaction is repulsive and the algorithm option \emph{addCannotLinkConstraints} is \texttt{True}, then the two clusters are constrained so that their members can never merge in subsequent steps of \emph{Phase~One}. If the interaction is attractive, then the clusters are merged, provided that they were not previously constrained. 
After each merge, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S, S')$ listed in Table \ref{tab:linkage-criteria}. \emph{Phase~One} terminates when all the remaining clusters are either constrained or share repulsive interactions. Note that, on unsigned graphs, in \emph{Phase 1} \algname{} merges all nodes in a single cluster and is then equivalent to a standard hierarchical clustering algorithm.

\emph{Phase Two:} Now that the clusters have grown in size, the algorithm removes the constraints previously introduced in \emph{Phase~One} and merges all the clusters that still share an attractive interaction, merging the most attractive one first\footnote{Note that in the version of \algname{} without \emph{cannotLinkConstraints}, nothing happens in \emph{Phase~Two} because all remaining interactions are repulsive.}. The final clustering $\Pi^*$ returned by \algname{} is found at the end of \emph{Phase~Two} and it is then composed of clusters sharing only mutual repulsive interactions.

Finally, in \emph{Phase~3}, the algorithm keeps merging all clusters until only a single one is left and then returns the hierarchical tree $T^*$ representing the full sequence of merging steps.
The algorithm was implemented using a standard HC implementation with computational complexity $\mathcal{O}(N^2 \log N)$ (details left in Appendix \ref{sec:detailed_impl}).

@@

\begin{algorithm}[t]
\footnotesize
  \begin{flushleft}
  \footnotesize
  \caption{\algname{}}
  % \caption{\algname{}: generalized algorithm for signed graph partitioning}
   \hspace*{\algorithmicindent} \textbf{Input:} Graph $\mathcal{G}(V,E,w^+,w^-)$; linkage criterion $\interact{}$; \\ 
   \hspace*{4.3em}boolean {\color{blue}addCannotLinkConstraints}  \\
  \hspace*{\algorithmicindent} \textbf{Output:} Final clustering $\Pi^*$, rooted binary hierarchical tree $T^*$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
  \footnotesize
  % \small
      \State Initial clustering: $\Pi=\{\{v_1\}, \ldots, \{v_{|V|}\}\}$
      \State Initialize hierarchical tree $T$ with leaf nodes  $V=\{v_1,\ldots,v_{|V|}\}$
      \State Initialize interactions between clusters with $\cost_e = w^+_e - w^-_e$
      \State \emph{// Phase 1: Merge positive interactions (possibly using constraints)}
      \State Push absolute interactions $|w_e|$ to priority queue $PQ$
      \Repeat 
        \State Pop $S,S'\in\Pi$ from $PQ$ with highest interaction $|\interact{}(S,S')|$
        \If{\big[{\color{green}\textbf{$\interact{}(S,S') > 0$}}\big] \textbf{and} \big[$S,S'$ \textbf{not} constrained\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T$
          \State Update interactions \& constraints with neighboring clusters
        \ElsIf{{\color{blue}addCannotLinkConstr} \textbf{and}  \big[{\color{red}\textbf{$\interact{}(S,S') \leq 0$}}\big]}
          \State Add CannotLink Constraint between $S$ and $S'$
        \EndIf
      \Until{\big[$PQ$ is empty\big]}
      % \State
      % \If{{\color{blue}addCannotLinkConstr}} % \Comment{Step 2: release constraints}
      \State \emph{// Phase 2: Remove constraints \& merge all positive interactions}
      \State Push signed interactions $\interact{}(S,S')$ to PQ, $\forall S, S' \in \Pi$
      \Repeat 
        \State Pop $S,S'\in\Pi$ from $PQ$ with highest interaction $\interact{}(S,S')$
        \If{\big[{\color{green}\textbf{$\interact{}(S,S') > 0$}}\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T$
          \State Update interactions with neighboring clusters
        \EndIf
      \Until{\big[{\color{red}\textbf{$\interact{}(S,S') \leq 0$}}\big]} 
      \State Save the final clustering $\Pi^* \gets \Pi$ 
      \State \emph{// Phase 3: Merge negative interactions until one single cluster is left}
      \Repeat
        \State Pop $S,S'\in\Pi$ from $PQ$ with highest interaction $\interact{}(S,S')$
        \State Merge clusters $S$, $S'$ and update hierarchical tree $T$
        \State Update interactions with neighboring clusters
      \Until{\big[Only one cluster is left in $\Pi$\big]} 
      % \EndIf
      \State
      \Return $\Pi^*$, $T^*\gets T$
  \end{algorithmic}
    \label{main_alg}
  \end{flushleft}

@@

\paragraph{Sum Linkage} -- 
On signed graphs, the sum of two attractive (or repulsive) interactions is still attractive (repulsive). On the other hand, on unsigned graphs, a strong attractive interaction could be obtained by summing many weak interactions, which depending on the application could be undesirable. This explains why, to our knowledge, an agglomerative algorithm with sum linkage has never been used on unsigned graphs. On signed graphs, such an algorithm was pioneered in \cite{levinkov2017comparative,keuper2015efficient} and was named Greedy Agglomerative Edge Contraction (GAEC)\footnote{An algorithm equivalent to GAEC was recently independently re-proposed in \cite{chehreghani2020hierarchical}.}. All variations of \algname{} decrease the multicut objective defined in Eq.~\ref{eq:MC_objective} each time two clusters with positive interaction are merged. But only GAEC always makes the \emph{greedy-choice} that mostly decreases the multicut objective at each iteration. The authors of \cite{levinkov2017comparative} also proposed a version of GAEC with cannot-link constraints named \emph{GreedyFixation}, but the constrained sum-link version of \algname{} proposed here (named \emph{HCC-Sum} in Table~\ref{tab:linkage-criteria}) did perform better in our experiments.

>>>>

\textbf{Linkage criteria and hierarchical trees} -- 
% We also denote as $S_u$ the cluster associated with node $u$ \TODO{Needed?}.
% We call two clusters $S,S'$ \emph{adjacent} if there exists at least one edge ${e_{ts}\in E}$ connecting a node $t\in S_u$ to a node $s\in S_v$. 
Let the interaction $\interact(S,S')\in\mathbb{R}$ between two clusters $S,S'$ be defined as a function, named \emph{linkage criterion}, depending on the weights of \emph{all} edges connecting clusters $S$ and $S'$, i.e. ${(S \times S') \cap E}$. 
The linkage criteria tested in this article are listed and defined in Table \ref{tab:linkage-criteria}.
A \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} representing the merging order of an agglomerative algorithm, such that the leaves of the tree are in one-to-one correspondence with $V$. 
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor $(u \vee v)\in T$ of nodes $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
% \TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
% Algorithm \ref{main_alg} returns a binary tree denoted by $T^*$.
Given an agglomerative algorithm with merging tree $T$, let $h_T:T \rightarrow \mathbb{N}$ denote the \emph{dendrogram-height} of each $(u\vee v)\in T$, which is defined as the iteration number at which nodes $u,v\in V$ were merged by the algorithm (see example in Fig.~\ref{fig:intro_figure}a). We also define $\treeHeight(u,v)$ as the signed interaction $\interact{}(S,S')$ between the two clusters $S,S'$ that were merged at iteration $h_T(u\vee v)$: 
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}\label{eq:def_dendr_interact}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T_{\mathrm{R}}[u \vee v]), \text{\texttt{leaves}}(T_{\mathrm{L}}[u \vee v]) \big)
\end{equation}
% \OPTIONAL{In other words, $h_T(u,v)$ and $\treeHeight(u,v)$ represent the \emph{dendrogram-height} and \emph{signed-interaction} associated to each $u \vee v$ node in the tree $T$ (see toy example in \TODO{Fig.~1}).}

@@

In phase one, \algname{} selects the pair of clusters with the highest absolute interaction $|\interact(S, S')|$, so that the most attractive and the most repulsive pairs are analyzed first. If the interaction is repulsive and the algorithm option \emph{addCannotLinkConstraints} is \texttt{True}, then the two clusters are constrained so that their members can never merge in subsequent steps of phase one. If the interaction is attractive, then the clusters are merged, provided that they were not previously constrained. 
After each merge, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S, S')$ listed in Table \ref{tab:linkage-criteria}. Phase one terminates when all the remaining clusters are either constrained or share repulsive interactions. Note that, on unsigned graphs, in phase one \algname{} merges all nodes in a single cluster and is then equivalent to a standard hierarchical clustering algorithm.

Phase two: Now that the clusters have grown in size, the algorithm removes the constraints previously introduced in phase one and merges all the clusters that still share an attractive interaction, merging the most attractive one first\footnote{Note that in the version of \algname{} without \emph{cannotLinkConstraints}, nothing happens in phase two because all remaining interactions are repulsive.}. The final clustering $\Pi^*$ returned by \algname{} is found at the end of phase two and it is then composed of clusters sharing only mutual repulsive interactions.

Finally, in phase three, the algorithm keeps merging all clusters until only a single one is left and then returns the hierarchical tree $T^*$ representing the full sequence of merging steps.
The algorithm was implemented using a standard HC implementation with computational complexity $\mathcal{O}(N^2 \log N)$ (details left in Appendix \ref{sec:detailed_impl}).

@@

\begin{algorithm}[t]
\footnotesize
  \begin{flushleft}
  \footnotesize
  \caption{\algname{}}
  % \caption{\algname{}: generalized algorithm for signed graph partitioning}
   \hspace*{\algorithmicindent} \textbf{Input:} Graph $\mathcal{G}(V,E,w^+,w^-)$; linkage criterion $\interact{}$; \\ 
   \hspace*{4.3em}boolean {\color{blue}addCannotLinkConstraints}  \\
  \hspace*{\algorithmicindent} \textbf{Output:} Final clustering $\Pi^*$, rooted binary hierarchical tree $T^*$\\
  \hspace*{\algorithmicindent} 
  \begin{algorithmic}[1]
  \footnotesize
  % \small
      \State Initial clustering: $\Pi=\{\{v_1\}, \ldots, \{v_{|V|}\}\}$
      \State Initialize hierarchical tree $T^*$ with leaf nodes  $V=\{v_1,\ldots,v_{|V|}\}$
      \State Initialize  cluster interactions with $\cost_e = w^+_e - w^-_e$, $\forall e\in E$
      \State \emph{// Phase 1: Merge positive interactions (possibly using constraints)}
      \State For all edges $e\in E$, push incident nodes of $e$ to PQ with priority $|w_e|$
      \Repeat 
        \State Pop $S,S'\in\Pi$ from $PQ$ with highest interaction $|\interact{}(S,S')|$
        \If{\big[{\color{green}\textbf{$\interact{}(S,S') > 0$}}\big] \textbf{and} \big[$S,S'$ \textbf{not} constrained\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
          \State Update interactions \& constraints with neighboring clusters
        \ElsIf{{\color{blue}addCannotLinkConstr} \textbf{and}  \big[{\color{red}\textbf{$\interact{}(S,S') \leq 0$}}\big]}
          \State Add CannotLink Constraint between $S$ and $S'$
        \EndIf
      \Until{\big[$PQ$ is empty\big]}
      % \State
      % \If{{\color{blue}addCannotLinkConstr}} % \Comment{Step 2: release constraints}
      \State \emph{// Phase 2: Remove constraints \& merge all positive interactions}
      \State Push signed interactions $\interact{}(S,S')$ to PQ, $\forall S, S' \in \Pi$
      \Repeat 
        \State Pop $S,S'\in\Pi$ from $PQ$ with highest interaction $\interact{}(S,S')$
        \If{\big[{\color{green}\textbf{$\interact{}(S,S') > 0$}}\big]}
          \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
          \State Update interactions with neighboring clusters
        \EndIf
      \Until{\big[{\color{red}\textbf{$\interact{}(S,S') \leq 0$}}\big]} 
      \State Save the final clustering $\Pi^* \gets \Pi$ 
      \State \emph{// Phase 3: Merge negative interactions until one single cluster is left}
      \Repeat
        \State Pop $S,S'\in\Pi$ from $PQ$ with highest interaction $\interact{}(S,S')$
        \State Merge clusters $S$, $S'$ and update hierarchical tree $T^*$
        \State Update interactions with neighboring clusters
      \Until{\big[Only one cluster is left in $\Pi$\big]} 
      % \EndIf
      \State
      \Return $\Pi^*$, $T^*$
  \end{algorithmic}
    \label{main_alg}
  \end{flushleft}

@@

\paragraph{Sum Linkage} -- 
On signed graphs, the sum of two attractive (or repulsive) interactions is still attractive (repulsive). On the other hand, on unsigned graphs, a strong attractive interaction could be obtained by summing many weak interactions, which depending on the application could be undesirable. This explains why, to our knowledge, an agglomerative algorithm with sum linkage has never been used on unsigned graphs. On signed graphs, such an algorithm was pioneered in \cite{levinkov2017comparative,keuper2015efficient} and was named Greedy Agglomerative Edge Contraction (GAEC)\footnote{An algorithm equivalent to GAEC was recently independently re-proposed in \cite{chehreghani2020hierarchical}.}. All variations of \algname{} decrease the multicut objective defined in Eq.~\ref{eq:MC_objective} each time two clusters with positive interaction are merged. But only GAEC always makes the \emph{greedy-choice} that mostly decreases the multicut objective at each iteration. The authors of \cite{levinkov2017comparative} propose an algorithm named \emph{GreedyFixation}, which, in our framework, is equivalent to phase one of \algname{} using cannot-link-constraints and a sum linkage. However, running both phase one and two of \algname{} with sum linkage (algorithm named HCC-Sum in this paper) performed better than GreedyFixation in our experiments, because some clusters are wrongly constrained in phase one and only merged in phase two. 
% removing constraints in phase two   However, in our experiments, we noted that constrained sum-link version of \algname{} proposed here (named \emph{HCC-Sum} in Table~\ref{tab:linkage-criteria}) .
